---
title: "Binomial distribution"
author: "Jörg Sahlmann"
date: "11.05.2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

## Binomial distribution

### Several ggplot-elements used

In the following figure the probabilities of x successes in n trials will be depicted as dots and lines. 

```{r}

x <- seq(0:50)
y <- dbinom(x, 50, 0.4)
y2 <- dbinom(x, 40, 0.6)

plot_data <- data.frame(x, y, y2)

ggplot(data = plot_data) +
  geom_point(aes(x = x, y = y)) + 
  geom_line(aes(x = x, y = y)) + 
  geom_segment(aes(x = x, xend = x, y = 0, yend = y), color = "blue", alpha = 0.5) +
  geom_point(aes(x = x, y = y2)) + 
  geom_line(aes(x = x, y = y2)) + 
  geom_segment(aes(x = x, xend = x, y = 0, yend = y2), color = "red", alpha = 0.5) +
  annotate("text", x = 40, y = 0.1, label = "Verteilung 1", color = "blue", hjust = 0) +
  ggtitle("Verteilungsbeispiel", subtitle = "Binomial") +
  xlab("Success") +
  ylab("Density")


```

### x successes in n trials

In the following figure the probabilities of x successes in n trials will be depicted as histogram. 

```{r}
# Idea from https://rpubs.com/mpfoley73/458411

plot_data <- data.frame(heads = 0:10, prob = dbinom(x = 0:10, size = 10, prob = 0.3))
plot_data$Heads <- ifelse(plot_data$heads == 2, "2", "other")

ggplot(data = plot_data, aes(x = factor(heads), y = prob, fill = Heads)) +
  geom_col() +
  geom_text(
    aes(label = round(prob,2), y = prob + 0.01),
    position = position_dodge(0.9),
    size = 3,
    vjust = 0
  ) +
  labs(title = "Probability of X = 2 successes.",
       subtitle = "B(X|n = 10, p = 0.3)",
       x = "Successes (x)",
       y = "Probability") 

```

### Zero to x successes in n trials

In the following figure the probabilities of x successes in n trials will be depicted as histogram. 

```{r}
# Idea from https://rpubs.com/mpfoley73/458411

plot_data <- data.frame(heads = 0:10, prob = dbinom(x = 0:10, size = 10, prob = 0.3))
plot_data$Heads <- ifelse(plot_data$heads <= 2, "<= 2", "other")

ggplot(data = plot_data, aes(x = factor(heads), y = prob, fill = Heads)) +
  geom_col() +
  geom_text(
    aes(label = round(prob,2), y = prob + 0.01),
    position = position_dodge(0.9),
    size = 3,
    vjust = 0
  ) +
  labs(title = "Probability of X <= 2 successes.",
       subtitle = "B(X|n = 10, p = 0.3)",
       x = "Successes (x)",
       y = "Probability") 

```

