---
title: "Normal distribution"
author: "Jörg Sahlmann"
date: "03.05.2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

## Probability Density function

Normal distribution with $\mu = 0$ and $\sigma = 1$.

$$\varphi (x) = \frac{e^{-\frac{x^2}{2}}}{\sqrt(2\pi)} $$


```{r }

x <- seq(-4, 4, by = 0.1)
y <- dnorm(x)
y2 <- pnorm(x)

plotdat <- tibble(x, y, y2)

ggplot(plotdat, aes(x = x, y = y)) +
  geom_line() +
  ggtitle("Probability Density Function", subtitle = "Normal distribution") +
  ylab("Density")

```

## Cumulative Density Function

Normal distribution with $\mu = 0$ and $\sigma = 1$.


$$ \Phi (x) = \frac{1}{\sqrt(2\pi)} \int_{-\infty}^{x} e^{-t^2/2} \ dt$$


```{r pressure}

ggplot(plotdat, aes(x = x, y = y2)) +
  geom_line() +
  ggtitle("Probability Density Function", subtitle = "Normal distribution") +
  ylab("Density")

```


## Probability Density function with tails

### One-sided


```{r}

tr <- qnorm(0.95)

ggplot(plotdat, aes(x = x, y = y)) +
  geom_area(data = plotdat %>% filter(x < tr), color = "black", fill = "lightgreen") +
  annotate("text", x = 0, y = 0.2, label = "95%") +
  geom_line() +
  ggtitle("Probability Density Function", subtitle = "Normal distribution") +
  ylab("Density")


```

### Two-sided


```{r}

tl <- qnorm(0.025)
tr <- qnorm(0.975)

ggplot(plotdat, aes(x = x, y = y)) +
  geom_area(data = plotdat %>% filter(x >= tl & x <= tr), color = "black", fill = "lightgreen") +
  annotate("text", x = 0, y = 0.2, label = "95%") +
  geom_line() +
  ggtitle("Probability Density Function", subtitle = "Normal distribution") +
  ylab("Density")


```


## Quantiles for probability density function

Which x-values represent 25%, 50% and 75% probability for the normal distribution?


```{r}

t1 <- qnorm(0.25)
t2 <- qnorm(0.50)
t3 <- qnorm(0.75)

c(t1, t2, t3)

ggplot(plotdat, aes(x = x, y = y)) +
  geom_area(data = plotdat %>% filter(x <= t1), color = "black", fill = "grey", alpha = 0.25) +
  annotate("text", x = t1, y = 0.1, label = "25%", hjust = 1) +
  geom_area(data = plotdat %>% filter(x <= t2), color = "black", fill = "grey", alpha = 0.25) +
  annotate("text", x = t2, y = 0.15, label = "50%", hjust = 1) +
  geom_area(data = plotdat %>% filter(x <= t3), color = "black", fill = "grey", alpha = 0.25) +
  annotate("text", x = t3, y = 0.2, label = "75%", hjust = 1) +
  geom_line() +
  ggtitle("Probability Density Function", subtitle = "Normal distribution") +
  ylab("Density")



```


## Rule of thumb

### Calculation

Rule of thumb for distribution of observations:

```{r}

sd1 <- pnorm(1) - pnorm(-1)
sd2 <- pnorm(2) - pnorm(-2)
sd3 <- pnorm(3) - pnorm(-3)


```

`r round(sd1*100, 1)`% of the observations are within $\pm 1$ standard deviation around the mean.

`r round(sd2*100, 1)`% of the observations are within $\pm 2$ standard deviation around the mean.

`r round(sd3*100, 1)`% of the observations are within $\pm 3$ standard deviation around the mean.


## Examples

### Example 1

Questionnaire scores are distributed nearly normally with mean 150 and standard deviation 30.

  - about 68.3% of the scores are between 120 and 180
  - about 95.4% of the scores are between 90 and 210
  - about 99.7% of the scores are between 60 and 240


### Example 2

To compare two results from different tests with different means and standard deviations the results have to be standardized.

$$ Z = \frac{observation - mean}{SD} $$

Test A scores are nearly distributed with mean 150 and standard deviation 30. Test B scores are nearly distributed with mean 22 and standard deviation 6.

Student 1 has a result of 170 on test A, student 2 has a result of 25 on test B.

Which result is further from the mean?

```{r}

r1 <- (170 - 150) / 30
r2 <- (25 - 22) / 6

ggplot(plotdat, aes(x = x, y = y)) +
  geom_line() +
  geom_vline(xintercept = r1, color = "red") +
  annotate("text", x = r1 + 0.1, y = 0.4, label = "Student 1", color = "red", hjust = 0) +
  geom_vline(xintercept = r2, color = "blue") +
  annotate("text", x = r2 + 0.1, y = 0.3, label = "Student 2", color = "blue", hjust = 0) +
  ggtitle("Probability Density Function", subtitle = "Normal distribution") +
  ylab("Density")



```

The result of student 1 is further from the mean.


### Example 3

A machine produces screws 400 mm long with a standard deviation of 9 mm. The length of the screws can be viewed as normally distributed.

#### Part 1

Calculate the probability that a screw is shorter than 385 mm.

```{r}

z <- (385 - 400) / 9
pz <- pnorm(z)
pz

```
The probability that a screw is shorter than 385 mm is `r round(pz*100, 1)`%.

#### Part 2

Calculate the probability that a screw is between 395 and 405 mm.

```{r}

zl <- (395 - 400) / 9
zr <- (405 - 400) / 9
pzl <- pnorm(zl)
pzr <- pnorm(zr)
pz <- pzr - pzl
pz

```

The probability that a screw is shorter between 395 and 405 mm is `r round(pz*100, 1)`%.

#### Part 3

Calculate the probability that the length of the screw deviates more than one standard deviation from the mean in either direction.

```{r}

pzl <- pnorm(-1)
pzr <- pnorm(1)
pz <- (1 - pzr) + pzl
pz

```


## Testing for normality

### Constructing a normal probability plot (QQ-plot)

```{r}

x <- c(3.41, 4.01, 5.08, 2.32, 6.46)
x <- sort(x)
pc <- seq(1, length(x))
pc <- pc / (length(x) + 1)
z <- qnorm(pc)

plotdat <- tibble(x, pc, z)

ggplot(plotdat, aes(x = z, y = x)) + 
  geom_point() + 
  geom_smooth(method = "lm", formula = y ~x)


```

### QQ-plot with provided plots

```{r}

ggplot(plotdat, aes(sample = x)) + 
  geom_qq() +
  geom_qq_line() +
  ggtitle("QQ-plot") +
  xlab("Theoretical") +
  ylab("Sample")

```




