---
title: "Tests for means"
author: "Jörg Sahlmann"
date: "04.05.2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

# One mean

## t-test

$H_0: \mu = \mu_0$


### Alternative 1

$H_{A1}: \mu > \mu_0$

Reject $H_0$ if the following condition is met:

$\hat{t} = \frac{\bar{x} - \mu_0}{s} \sqrt{n} > t_{n-1; 1-\alpha}$

p is the area under the $t_{n-1}$ distribution right of $\hat{t}$.

#### Example

The temperature in an incubator shall have more than 37°C. The following temperatures have been measured:

36.9 37.3 37.6 37.1 37.0 37.5 37.5 37.9

```{r}

t_temp <- c(36.9, 37.3, 37.6, 37.1, 37.0, 37.5, 37.5, 37.9)
t.test(t_temp, alternative = "greater", mu = 37)

```

The temperature is significantly higher than 37°C.


### Alternative 2

$H_{A2}: \mu < \mu_0$


### Alternative 3

$H_{A3}: \mu \ne \mu_0$


```{r}


```


# Two means


```{r}


```


# More than two means


```{r}


```

