---
title: "Tests for proportions"
author: "Jörg Sahlmann"
date: "4 5 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

# Definitions

$p$ is the proportion in the population.

$\hat{p}$ is the proportion in the sample.

For the confidence interval $\hat{p}$ will be used for the standard error.

For the hypothesis test $p$ will be used for the standard error.


# One proportion

## Confidence interval for one proportion

$\hat{p} \pm z^*SE_\hat{p}$

Margin of error: $ME = z^*SE_{\hat{p}}$

### Example 1

A question was answered correctly by 570 out of 671 students from a university with 10,000 students.

Give a confidence interval for the proportion of the correct answer from all students.

```{r}

phat <- 570/671

cil <- phat + qnorm(0.025) * sqrt(phat * (1 - phat) / 671)
cir <- phat + qnorm(0.975) * sqrt(phat * (1 - phat) / 671)
me <- 1.96 * 0.0138
me
cil <- 0.85 - me
cir <- 0.85 + me
cil
cir

prop.test(570, 671, p = NULL, alternative = "two.sided", correct = FALSE)

```

### Example 2

How many students do we need if we want to narrow the margin of error from 0.0270 to 0.0100?

We use the formula for the confidence interval and solve it for n.

  - ME = 1.96 * sqrt(phat * (1 - phat) / n)

  - 0.01 = 1.96 * sqrt(phat * (1 - phat) / n)

  - (0.01/1.96)^2 * n = phat * (1 - phat)

  - n = phat * (1 - phat) / (0.01/1.96)**2


```{r}

phat = 570/671
n = phat * (1 - phat) / (0.01/1.96)**2
n

```

We need at least 4913 students to have a margin of error of maximal 0.01.

Solving the equation above can also be done with wxmaxima and the following code snippet.




## Hypothesis testing for one proportion

1. Set hypothesis according to research question:

$H_0: p = null \ value$
$H_{A_1}: p < null \ value$
$H_{A_2}: p > null \ value$
$H_{A_3}: p \ne null \ value$

2. Calculate the point estimate: $\hat{p}$

3. Check conditions:

  - Independence: Sampled observations must be independent (random sample/assignment and if sampling without replacement, n < 10% of population)
  - Sample size: np >= 10 and n(1-p) >= 10

4. Draw sampling distribution, shade p-value, calculate test statistic:

$Z = \frac{\hat{p} - p}{SE}$, $SE = \sqrt{\frac{p (1-p)}{n}}$

5. Make a decision, and interpret it in the context of the research question


# Two proportions

## Difference of proportions

Parameter of interest is the difference of proportions between all of respective elements from the population.

Point estimate is the difference of proportions from the sample.

point estatimate $\pm$ margin of error

$(\hat{p_1} - \hat{p_2}) \pm z^*SE_{\hat{p_1} - \hat{p_2}}$

$SE = \sqrt{\frac{\hat{p_1} (1 - \hat{p_1})}{n_1} + \frac{\hat{p_2} (1 - \hat{p_2})}{n_2}}$


### Example 1

One group of students and one group of hunters was asked about shooting as private sport.

750 of 950 hunters said yes, 110 of 550 students said yes. Is there a signifcant difference?

```{r}

n1 <- 950
n2 <- 550
p1 <- 750/n1
p2 <- 110/n2

cil <- (p1 - p2) - 1.96 * sqrt((p1 * (1 - p1)/n1) + (p2 * (1 - p2)/n2))
cir <- (p1 - p2) + 1.96 * sqrt((p1 * (1 - p1)/n1) + (p2 * (1 - p2)/n2))

cil
cir

prop.test(x = c(750, 110), n = c(950, 550))

```

The difference of the proportions differs significantly from zero.



# More than two proportions


```{r}


```

